Using simulation to optimize the sampling design of fisheries-independent trawl surveys

### 2018-09-11 analysis/paper/sim_survey_paper.Rmd
Add a RMSE comparison table to the paper

@@

A dynamic population that varies across ages, time and space is simulated with the parameter settings used here (Table \@ref(tab:par)). Figure \@ref(fig:sp-fig) depicts the aggregate distribution of the population in one year of the simulation and the corresponding catches of one stratified random survey over this population. Layered within this, of course, is the distributions of each age, as well as age and length sampling details. Though age-specific details are not depicted for simplicity, the general pattern of spatial aggregation apparent in Figure \@ref(fig:sp-fig) is similar across all ages. That is, all ages tend to aggregate in fairly dense clusters and low-density zones are fairly wide ranging. The survey, therefore, tends to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:sp-fig)). The samples within each successful set are also dynamic and rarely include all age groups. This is because only moderate correlation was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations. This results in an intra-haul correlation of approximately `r sim_icc`, which, in this case, is of equal magnitude to the intra-haul correlation observed across more than 20 years of survey data of 3Ps cod.

@@

|      **Parameter name** |      **Symbol** |                                       **Setting** |
|:----------------------- | :-------------- | :------------------------------------------------ |
|             Set density |      $D_{sets}$ |    0.0005, 0.001, 0.002, 0.005, 0.01 sets / km^2^ |
|  Length sampling effort |   $M_{lengths}$ |       5, 10, 20, 50, 100, 500, 1000 lengths / set |
|     Age sampling effort |      $M_{ages}$ |   2, 5, 10, 20, 50 ages / length group / division |

@@

```{r fig_fun}

```{r sp-fig, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^. Grey lines indicate depth-based survey strata."}

@@

```{r tot-fig, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$. The heavy dark line indicates the true tred in the total population available to the survey."}

@@

```{r age-fig, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$. Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios. The heavy dark line indicates the true population available to the survey."}

@@

```{r year-fig, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$. Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios. The heavy dark line indicates the true trend in the population available to the survey."}

@@

```{r error-fig, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol. Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)."}

@@

```{r appen-age-fig, cache = TRUE, fig.width = 8, fig.height = 6}

@@

```{r appen-year-fig, cache = TRUE, fig.width = 8, fig.height = 6}

@@

```{r appen-error-fig, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide"}

>>>>

```{r fun}

@@

A dynamic population that varies across ages, time and space is simulated with the parameter settings used here (Table \@ref(tab:par)). Figure \@ref(fig:sp) depicts the aggregate distribution of the population in one year of the simulation and the corresponding catches of one stratified random survey over this population. Layered within this, of course, is the distributions of each age, as well as age and length sampling details. Though age-specific details are not depicted for simplicity, the general pattern of spatial aggregation apparent in Figure \@ref(fig:sp) is similar across all ages. That is, all ages tend to aggregate in fairly dense clusters and low-density zones are fairly wide ranging. The survey, therefore, tends to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:sp)). The samples within each successful set are also dynamic and rarely include all age groups. This is because only moderate correlation was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations. This results in an intra-haul correlation of approximately `r sim_icc`, which, in this case, is of equal magnitude to the intra-haul correlation observed across more than 20 years of survey data of 3Ps cod.

@@

|      **Parameter name** |      **Symbol** |                                       **Setting** |
|:----------------------- | :-------------- | :------------------------------------------------ |
|             Set density |      $D_{sets}$ |    0.0005, 0.001, 0.002, 0.005, 0.01 sets / km^2^ |
|    Length sampling rule |   $M_{lengths}$ |       5, 10, 20, 50, 100, 500, 1000 lengths / set |
|       Age sampling rule |      $M_{ages}$ |   2, 5, 10, 20, 50 ages / length group / division |

Table: (\#tab:comp) - Root-mean-squared errors (RMSE) and number of samples [sets ($N_{sets}$), lengths ($N_{lengths}$) and ages ($N_{ages}$)] obtained from an array of surveys with different sampling protocol [set densities ($D_{sets}$), maximum sampling effort of lengths per set ($M_{lengths}$), and maximum number of ages to collect per length group ($M_{ages}$)]. Differences from the minimum RMSE value ($\Delta ~ RMSE$) within each set density option are presented to aid comparisons.

```{r comp, results = "asis"}

totals <- sim$samp_totals[, list(n_sets = mean(n_sets), n_caught = mean(n_caught),
                                 n_measured = mean(n_measured), n_aged = mean(n_aged)),
                          by = "survey"]
errors <- merge(sim$surveys, sim$age_strat_error_stats, by = "survey")
d <- merge(errors, totals, by = "survey")

which_set_dens <- c(0.0005, 0.002, 0.01)
which_len_caps <- c(10, 100, 1000)
which_age_caps <- c(5, 10, 50)

tab <- d[d$set_den %in% which_set_dens & d$lengths_cap %in% which_len_caps & 
           d$ages_cap %in% which_age_caps, 
         list(lengths_cap, ages_cap,
              n_sets = round(n_sets), n_measured = round(n_measured), n_aged = round(n_aged),
              RMSE = RMSE, delta_RMSE = RMSE - min(RMSE)), 
         by = "set_den"]
tab <- tab[order(set_den, delta_RMSE), ]
tab$set_den <- formatC(tab$set_den)
tab$set_den[tab$delta_RMSE != 0] <- NA
tab$RMSE <- si(tab$RMSE)
tab$delta_RMSE <- si(tab$delta_RMSE)

split_tab <- split(tab, tab$n_sets)
for (i in seq_along(split_tab)[-length(split_tab)]) {
  split_tab[[i]] <- rbind(split_tab[[i]], data.table(RMSE = NA), fill = TRUE)
}
tab <- rbindlist(split_tab)

names(tab) <- c("$D_{sets}$", "$M_{lengths}$", "$M_{ages}$", 
                "$N_{sets}$", "$N_{lengths}$", "$N_{ages}$",
                "$RMSE$", "$\\Delta ~ RMSE$")

options(knitr.kable.NA = "")
knitr::kable(tab, align = "r", format = "markdown")

@@

```{r sp, cache = TRUE, fig.height = 3.8, fig.width = 8.5, fig.cap = "- An example from year seven of the simulation of a) the distribution of total abundance available to the survey, and b) locations of the sets conducted and total numbers caught in a stratified random survey with a set density of 0.002 sets / km^2^. Grey lines indicate depth-based survey strata."}

@@

```{r tot, cache = TRUE, fig.width = 8, fig.height = 3, fig.cap = "- Fanchart of stratified estimates of the trend in total abundance from surveys with different set densities, $D_{sets}$. The heavy dark line indicates the true tred in the total population available to the survey."}

@@

```{r age, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of abundance at age from year seven of the simulation from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$. Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios. The heavy dark line indicates the true population available to the survey."}

@@

```{r year, cache = TRUE, fig.width = 8, fig.height = 6, fig.cap = "- Fanchart of stratified estimates of the trend in abundance at age four from surveys with different set densities, $D_{sets}$, and length sampling protocol, $M_{lengths}$. Number of ages sampled per length group, $M_{ages}$, was 10 in all scenarios. The heavy dark line indicates the true trend in the population available to the survey."}

@@

```{r error, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide", fig.cap = "- Wireframe plot of root-mean-squared error (RMSE) from an array of surveys with different sampling protocol. Panels represent surveys with different set densities ($D_{sets}$), x-axes represent the maximum sampling effort of lengths per set ($M_{lengths}$), and y-axes represent the maximum number of ages to collect per length group ($M_{ages}$)."}

@@

```{r appen-age, cache = TRUE, fig.width = 8, fig.height = 6}

@@

```{r appen-year, cache = TRUE, fig.width = 8, fig.height = 6}

@@

```{r appen-error, cache = TRUE, fig.height = 3, fig.width = 8.5, results = "hide"}
